### 2019-03-29 paper/sex_selection_spacing-ver2.tex
Finish shortening paper to < 8,000 words

Average spacing is a convenient way to understand the overall changes in behavior but may 
hide important differences in the distribution of spacing.
Specifically, whether increases in average spacing come from fewer births after very short 
intervals, from even longer spacing of births with already long spacing, or from a general 
increase in spacing have important implications for what effects the increases in average 
spacing are likely to have and how to interpret the motivation for the changes.
This section, therefore, examines how spacing is distributed within spells.
I show survival curves conditional on predicted parity progression rather than standard 
survival curves, which allows direct comparisons of the distribution of spacing across 
groups, independent of differences in parity progression probabilities.

@@

Figure \ref{fig:pps_low} shows the distribution of birth within a spell for women with no 
education.
For the second spell, what is most interesting is that there is almost no difference 
between those with a first-born boy and those with a first-born girl.
For the 1972--1984 period the two conditional survival curves only begin to diverge 
at around 18 months (corresponding to 27 months after the first birth took place), and
at that point close to 50 percent of the women have already had their second child.
With a nine-month pregnancy, this means that almost 50 percent of the children born are
born within the interpregnancy interval associated with a higher risk of adverse health
outcomes and mortality, and, somewhat surprisingly, that there do not appear to be any 
particular disadvantage for the child born after a girl than after a boy.%
\footnote{
Note that this does not imply that the first-born girl is not in a more adverse position
than a first-born boy when the second child arrives, just that this is not directly
related to a short birth interval.
}
The difference in spacing after a first-born girl and a first-born boy remains almost 
the same across the four decades.
Furthermore, the proportion with very short spacing remains almost the same.

For the third spell, the son preference driven short spacing is evident throughout the
distribution and becomes more pronounced over the four decades.
Families with two girls are especially likely to have a birth early with 25 percent
having a child within 12 months of the beginning of the spell (21 months after the second 
birth).
That number is only slightly smaller four decades later. 
The fourth spell shows changes in the distribution that appear to be more pronounced
than for the third spell, although these results are based on fewer observed births.
For families with three girls, 75 percent have had their fourth child within 36 months 
after the beginning of the spell, but this occurs more than a year later for families
with two or more boys.

@@

The distribution of spacing for the second spell is almost indistinguishable 
between urban women with eight or more years of education and rural women 
with no education for the 1972--1984 period, as shown in Figure \ref{fig:pps_high}.
The small differences point to these two very different groups of women behaving similarly
in response to son preference in a situation where there was little access to
prenatal sex determination or little incentive to use it early even for those 
who might have had access.
However, whereas the no education group shows almost no change across the four decades for 
the second spell, the high education women show an upward shift and a relatively even
distribution of births within the spell for both women with a first-born son and women
with a first-born girl.
The result is that the spacing patterns are close to identical after first-born 
girls and after first-born boys until after approximately 80 percent have had their 
second child, although as mentioned above, this is driven in large part by the significant
use of sex selection among women with a first-born girl.

For the third spell, the 1972--1984 period shows a close to standard pattern throughout, 
although, as mentioned above, the sex ratios for women without girls was substantially 
higher than the natural sex ratio either because of recall error or early use of 
sex selection.
The correlation between the increased use of sex selection and the reversal of the 
standard spacing pattern show up clearly in the 2005--2016 period for the third spell with 
spacing to the third child consistently longer with two girls than with either two boys 
or one boy and one girl throughout the entire spell.
This pattern is most consistent with early and continued use of sex selection, rather 
than a general reduction in the likelihood of very short birth intervals.
It also means that the previously born girls can expect more time without a younger sibling,
which may reduce the sibling competition for resources.

There is an even more pronounced rightward and upward shift for spacing after only
girls in the fourth spell for the first years of the spell.%
\footnote{
The caveat is that there are few births to base the results on,
especially for women with two or more sons.
} 
Furthermore, the differences are substantial across most of the distribution.
Twelve months after the beginning of the spell there is a more than ten 
percentage points difference in the conditional survival curves for only girls
and only boys.
Both of the spells, therefore, show changes in the distribution of births that is in line
with extensive use of sex selection among families with no or one son.

@@

To understand how regional differences in son preference, fertility, and use of sex 
selection affect birth spacing, I divide India into four broad regions based on the degree 
of son preference and initial fertility levels.
The focus is on the most educated women since they are the primary users of sex selection.
The four regions are based on \citet{retherford03b}, although expanded to include all 
states in the surveys.
Table \ref{tab:regions} shows the four regions and the states that they contain.%
\footnote{
The geographical names do not entirely reflect the geography of India, and mainly
serves as mnemonic short-hands.
}
The state names are those in existence at the time of the NFHS-1 data collection.
States formed later are allocated as close as possible to the original state.
The ``West'' contains states with a high degree of son preference and use of sex selection.
The ``North'' consists of states that also have a strong son preference, but initially had
higher fertility and, therefore, lower likelihood of using sex selection.
The ``East'' states generally have only a moderate level of son preference.
Finally, the ``South'' contains states with traditionally low son preference and lower
fertility than the rest of India.

@@

@@

Figure \ref{fig:bootstrap_high_r1} shows the predicted average birth intervals, sex 
ratios, and likelihoods of having a child by the end of the spell over time for the three 
spells in the ``West.''
What most stands out is how close to the national estimates the ``West'' estimates are.
The predicted second spell average birth spacing is similar across families with a girl
or a boy, and the average spacing substantially higher in the absence of boys than when
there is at least one boy for the third spell.
The main difference is that the very high predicted sex ratios reflect the expected higher 
use of sex selection.
For the second spell, the predicted sex ratios are around 60 percent for both urban and
rural women, and for the third spell it is around 70 percent for urban women and 
consistently above 60 percent for rural women.
This higher use of sex selection, in turn, translates into longer average spacing.
Furthermore, fertility has fallen faster than the national trend, especially for women
with one or more sons.

@@

Figure \ref{fig:bootstrap_high_r2} shows the same outcomes for the ``North'' region. 
Despite the idea that the ``North'' has high fertility and lower 
likelihood of sex selection the differences to the ``West'' region are surprisingly small,
at least in the urban areas.
Fertility is, indeed, higher than in the ``West'' region, but the declines in the 
probability of parity progression have been relatively rapid and mirror the national one.
The use of sex selection is also close to the ``West,'' and there is the same reversal of
spacing pattern as there.
The rural areas of the ``North'' are a slightly different story.
Fertility is substantially higher, and the use of sex selection much lower than in the
rural areas of the ``West.'' 
That said, there still is clear evidence that the sex ratio is substantially above normal
at close to 60 percent for the latest decade.

@@

The results for the ``East'' region of India are shown in Figure 
\ref{fig:bootstrap_high_r3}, although it does not include the fourth spell results because 
the sample size is smaller for this region and the fourth spell results very noisy.
Despite the ``South'' region generally considered to be the low fertility region, the
reduction in fertility has been even faster in the ``East.''
For the last decade, only half of the urban women with a first-born son are predicted to 
have a second child by the end of the spell, which means almost nine years after the first 
birth.
The probability is only slightly higher for those in urban areas with a first-born 
daughter and still less than 70 percent.
Not surprisingly, the probabilities are higher in rural areas, but still below the 
national pattern.
Similarly, conditionally on having a second child, the probabilities of having a third
child are low, even in the absence of sons.
In urban areas, a woman with two girls has only around a 40 percent likelihood of 
having a third child.

What stands out, however, is the rapid increase in the lengths of birth intervals,
especially for the second spell.
In both urban and rural areas, the predicted average birth intervals have gone up by
more than a year across the four decades, and that is independent of whether the first-born
was a boy or a girl.
The increases in birth intervals are slightly smaller for the third spell but still 
substantial.
Although the spacing pattern is consistent with some son preference, with the average
spacing slightly shorter in the absence of boys than when at least one boy is present,
there is little evidence of sex-selective abortions.
The sex ratios after a first-born daughter are slightly elevated but much lower than 
the prior two regions.
There is more evidence for the third spell, where the sex ratio is close to 60 percent
in the last period for urban women and the last two for rural women, but recall that 
there is a much smaller likelihood of even making it to the third spell for this region
than the other or the national average.

@@

The final region is the ``South,'' shown in Figure \ref{fig:bootstrap_high_r4}.
This region is historically thought to have the lowest degree of son preference of the
regions of India.
As for the ``East,'' I do not show the fourth spell results because of the small sample 
size and the associated noisy estimates.
Judging from the second spell progression rate and the associated spacing pattern, it
does, initially, appear that there is little son preference in this region.
The likelihood of having a second birth is relatively high at around 80 to 90 percent
and is close to independent of the sex of the first birth in both urban and rural areas.
The predicted average spacing also show no sign of son preference, and the predicted
sex ratios are close to the natural rate.
The average birth spacing started high relative to the other regions, and with little 
change in fertility and no significant use of sex selection, it has increased only 
slightly over time.

Despite the second spell's apparent lack of son preference behavior, for the third spell 
in urban areas the average birth interval has increased substantially for those with only 
girls relative to those with a boy and a girl, corresponding to
a predicted sex ratio close to 60 percent with only girls. 
The signs of the use of sex selection also show up for rural women with no sons.
Furthermore, over the four decades, the likelihood of having a third birth has decreased
much less than for either a boy and a girl or two boys.%
\footnote{
The relatively high parity progression for women without son does raise the question of
whether ``daughter avoidance'' is behind the fertility behavior in this region as
suggested by \citet{DiamondSmith2008}.
}
The differences in parity progression and the sex ratios suggest that despite originally
having less son preference, the ``South'' now appears to have longer average spacing 
in the absence of sons and the higher sex ratio associated with the use of sex selection.

@@

The central question addressed here is the extent to which spacing patterns significantly 
changed as prenatal sex determination became available.
The underlying idea is that in the absence of sex selection son preference leads to shorter 
spacing after the birth of a girl than after the birth of a boy, whereas when prenatal sex 
selection is available son preference increases the spacing after the birth of a girl 
relative to after the birth of a boy.
I introduce an empirical method that simultaneously accounts for spacing between births 
and the potential use of sex selection. 
I apply the method to over four decades of data from India's NFHS.

Three major, interrelated trends over the four decades emerge from the data.
First, fertility has declined for all groups, as shown by the parity progression
probabilities, with a larger decline the more educated the woman.
Across all groups, the likelihood of having an additional child in each spell depends 
strongly on the number of sons, with women with none or only one son having
the highest probability of a birth.
Second, there has been a general increase in the length of average birth 
intervals for all education groups.
The increase in the average birth interval is larger the higher the spell and the higher the 
education group.
Finally, as shown by the increasing sex ratios for women with no or only one son there
has been a substantial increase in the use of sex selection over time.
As for fertility and birth intervals, this increase is larger the higher the spell and
with higher education.

The results show two very different approaches to son preference with significant
impacts on spacing patterns.
At one extreme, rural women without education mostly follow the standard pattern of 
shorter spacing when a woman does not have the desired number of sons.
The strength of this pattern does, however, depend critically on spell number.
For the second spell, although the average interval after a first-born girl is shorter
than the average interval after a first-born boy, the difference is small.
The differences in spacing are larger for the third spell and largest for the fourth.
For the fourth spell, the divergence in intervals across sex compositions has 
increased over time, with the largest increases in average intervals for women with 
two or three boys.
There has also been an increase in average spacing for women with only girls, and there
is some limited evidence of biased sex ratios, which suggests that this increases in
spacing follow from the use of sex selection.

At the other extreme, urban women with eight or more years of education have experienced 
an almost complete reversal of the traditional spacing patterns as a result of falling
fertility and the associated extensive use of sex selection.
Women with either no or one son now have substantially longer spacing than if they have 
two or more sons, rather than having the next birth sooner as they mostly did before sex 
selection became available.
As shown by the increasingly significant deviations in the sex ratio from the natural sex 
ratio, these changes are due to increased use of sex selection rather than a sudden
reversal in son preference.
Only the second spell has not seen a reversal despite a substantially increased use of
sex selection, mainly because the average spacing with a first-born boy has increased 
independently by more than nine months.

Despite the differences in average spacing across sex compositions of prior children,
the picture of how son preference affects spacing is more complicated than excepted.
First, for the second spell, the differences in birth intervals for the first 25 percent of 
births are negligible across sex compositions, despite that these birth intervals are
clearly within the durations considered the most likely to generate adverse outcomes.
Furthermore, there is little difference in this spacing patterns across education.
Hence, most of the divergence in spacing patterns across sex compositions arise from
longer spells, which are likely to have less negative impacts on children and the mother.
Second, one of the side-effects of the extensive use of sex selection for third and fourth 
spells is that only-girls families are much less likely to have a short, i.e., less than 
18 months, interpregnancy interval, which may reduce the sibling competition for resources.

With a country as large as India, it is natural to expect variation in both fertility
behavior, spacing patterns, and the degree of son preference.
The regional results confirm this, although differences are less apparent than one would
expect given the prior research on regional differences.
The ``North'' and the ``West'' regions have relatively similar spacing patterns, likely
precipitated by rapidly declining fertility and substantial use of sex selection.
The exception is that the rural areas of the ``North'' still have higher fertility and
consequently less use of sex selection and have, therefore, not yet experienced the
same reversal of spacing patterns.
Even the ``South,'' which is otherwise considered the area with the least son preference,
has seen a substantial increase in the use of sex selection and the associated changes 
in the spacing patterns.
Even though the ``South'' is often mentioned as the low-fertility region of India, 
fertility has declined fastest in the ``East'' region, although this does not appear to 
have translated into the use of sex selection.

The results here lead to a set of broader questions that future research should address.
First, what is behind the substantial increases in spacing even for women who are unlikely 
to use sex selection, and, related, why does spacing appear to increase with declining 
parity progression probabilities?
The distributional results provide an initial indication that most of the increases in 
average birth intervals come from the upper end of the birth intervals rather a general
increase across the distribution.
It is, however, unclear why we are seeing relatively more very long birth intervals over
time, and the longer average spacing seems to run counter to a general reduction in the 
length of women's reproductive spans found in earlier research \citep{Padmadas2004}.
Understanding whether the longer average spacing is by choice at the beginning of the 
spell, because of unplanned pregnancies, or because of changes in economic circumstances 
that allow families to have an additional child later has important implications for both 
policy design and outcomes for mother and children.

Second, to what extent are the improvements in health for girls relative to boys the result 
of selection, the longer spacing between births, or changing son preference?
We know that health outcomes for girls appear to improve in the presence of sex selection 
\citep{Lin2014,Hu2015}.
Furthermore, there is some evidence that sample selection can make it appear that girls 
are healthier, even though the underlying cause is a combination of sex selection and 
higher mortality together with recall error \citep{Portner2018a}. 
Understanding what role these factors play in better health for girls is especially 
important when evaluating policies that aim to directly limit the use of sex selection 
rather than changing preferences or incentives.
If the better health outcomes for girls are, for example, an unintended side-effect of the 
longer spacing that arises from sex-selective abortions, rather than because the smaller 
number of girls makes them more valued as is often assumed, then an effective ban on sex 
selection may, at least temporarily, worsen health outcomes for girls.

Finally, we need to understand how female labor force participation interacts with the use 
of sex selection.
Increased autonomy for women, arising, for example, from better opportunities for working 
outside the home, has been suggested as a way to increase women's status and thereby lower 
the use of sex selection \citep{Das-Gupta2016}.
This approach may, however, be a double-edged sword.
On the one hand, it is a clear benefit to the women who gain bargaining power, and it 
increases the cost of repeated sex-selective abortions because the increased duration 
between births would cause a stronger disruption in labor market participation.
On the other hand, it may further lower desired fertility, and that may, everything else 
equal, lead to higher use of sex selection.
Understanding the trade-off between long-term benefits from improvements in women's labor 
force participation and short-term costs from potential increases in sex selection is of 
paramount importance.

>>>>

Average spacing is a convenient way to understand the overall changes in behavior but may 
hide crucial differences in the distribution of spacing.
Specifically, whether increases in average spacing come from fewer births after very short 
intervals, from an even longer spacing of births with already long spacing, or a general 
increase in spacing have important implications for what effects the increases in average 
spacing are likely to have and how to interpret the motivation for the changes.
This section, therefore, examines the distribution of spacing within spells.
I show survival curves conditional on predicted parity progression rather than standard 
survival curves, which allows direct comparisons of the distribution of spacing across 
groups, independent of differences in parity progression probabilities.

@@

The most striking results for women with no education are that, in both periods, close to 
50\% had their second child by 18 months and that there are minimal differences in the 
probability of very short intervals whether the first-born was a boy or a girl.
Hence, almost 50\% of second-borns are born within an interval associated with a higher 
risk of adverse outcomes, and, surprisingly, there do not appear to be any particular 
disadvantage for the child born after a girl than after a boy.
There is more evidence of son-preference driven short spacing for the third and fourth
spells, but a large number of women still have very short intervals whether they have boys 
or girls already.
For both spells, the difference in average spacing mainly come from the upper end of the
distribution; for example, the 75th percentile show a twelve months difference between 
families with three girls and families with two or more boys.

@@

Another remarkable result is the similarity in the lower end of the distribution of
spacing for the second spell between urban women with eight or more years of education and 
rural women with no education for the 1972--1984 period with little difference across
sex composition and close to 50\% with very short birth intervals.
However, the high education women show an upward shift over time and spacing patterns that
are close to identical after first-born girls and after first-born boys until after 
approximately 80 percent have had their second child, driven by the significant use of sex 
selection among women with a first-born girl.

The sex selection induced reversal of the standard spacing pattern shows up clearly in the 
2005--2016 period for the third and fourth spells, with spacing to the next child 
consistently longer with only girls than with the other sex compositions.
This pattern is most consistent with early and continued use of sex selection, rather 
than a general reduction in the likelihood of very short birth intervals.
It also means that the previously born girls can expect more time without a younger 
sibling, which may reduce the sibling competition for resources.

% Furthermore, the differences are substantial across most of the distribution.
% Twelve months after the beginning of the spell there is a more than ten 
% percentage points difference in the conditional survival curves for only girls
% and only boys.
% Both of the spells, therefore, show changes in the distribution of births that is in line
% with extensive use of sex selection among families with no or one son.

@@

To understand how regional differences in son preference, fertility, and use of sex 
selection affect birth spacing, I divide India into four regions, shown in Table 
\ref{tab:regions}, based on the degree  of son preference and initial fertility 
levels \citep{retherford03b}.%
\footnote{
The state names are those used in NFHS-1, with states formed later allocated as close as 
possible to the original state.
}
The ``West'' contains states with a high degree of son preference and use of sex selection.
The ``North'' consists of states that also have a strong son preference, but initially had
higher fertility and, therefore, lower likelihood of using sex selection.
The ``East'' states generally have only a moderate level of son preference.
Finally, the ``South'' contains states with traditionally low son preference and lower
fertility than the rest of India.

@@

I focus on the most educated women as the primary users of sex selection.
Figures \ref{fig:bootstrap_high_r1}, \ref{fig:bootstrap_high_r2}, 
\ref{fig:bootstrap_high_r3}, and \ref{fig:bootstrap_high_r4} show the predicted average 
birth intervals, sex ratios, and parity progression probabilities for the ``West,''
``North,'' ``East,'' and ``South,'' respectively.
The ``East'' and ``South'' do not include the fourth spell because of small sample sizes 
and very noisy results.

@@

Despite the idea that the ``West'' and ``North'' differ in fertility and likelihood of 
using sex selection, the differences are surprisingly small, at least for urban areas.
Furthermore, both are close to the national averages presented above, with the main 
difference that the very high sex ratios reflect a higher use of sex selection, 
presumably as a result of fertility falling faster than the national trend.
The predicted sex ratios in urban areas are around 60\% for the second spell and
around 70\% for the third spell.
This higher use of sex selection, in turn, translates into the reversal of standard
spacing pattern.

The two regions are further apart for the rural areas.
In the ``West'' the rural areas show only slightly lower use of sex selection and 
therefore a less pronounced reversal in spacing patterns than in the urban areas.
In the rural ``North,'' fertility is substantially higher, and the use of sex selection 
subsequently lower, but the sex ratio is still substantially above normal at close to 60 
percent for the latest decade.

@@

@@

What stands out for the ``East'' is the rapid increase in the lengths of birth intervals,
especially for the second spell, where the increases have been more
than a year, independently of whether the first-born was a boy or a girl.
The increases in birth intervals are slightly smaller for the third spell but still 
substantial.
Although the spacing pattern is consistent with some son preference, there is little 
evidence of sex-selective abortions.
Furthermore, despite the ``South'' is generally considered the low fertility region, 
the reduction in fertility has been even faster in the ``East.''
For example, only half of the urban women with a first-born son are now predicted to 
have a second child by the end of the spell.

@@

The ``South'' is historically thought to have the lowest degree of son preference,
which has initial support from the second spell's progression rates, spacing pattern, and 
sex ratios.
However, for the third spell in urban areas, the average birth interval has increased 
substantially, and the predicted sex ratio is close to 60\% for those with only girls. 
The signs of the use of sex selection also show up for rural women with no sons.
Furthermore, the likelihood of having a third birth has decreased much less with no sons 
than the other sex compositions.%
\footnote{
The relatively high parity progression for women without son runs counter to what we
should expect if ``daughter avoidance'' is behind the fertility behavior as
suggested by \citet{DiamondSmith2008}.
}
Hence, the ``South'' now also shows the longer average spacing in the absence of sons and 
the higher sex ratio associated with the use of sex selection.

@@

The central question addressed here is the extent to which spacing patterns significantly 
changed as prenatal sex determination became available.
The underlying idea is that in the absence of sex selection son preference leads to shorter 
spacing after the birth of a girl than after the birth of a boy, whereas when prenatal sex 
selection is available son preference increases the spacing after the birth of a girl 
relative to after the birth of a boy.
I introduce an empirical method that simultaneously accounts for spacing between births 
and the potential use of sex selection, and apply the method to four decades of data 
from India's NFHS.

Three major, interrelated trends emerge.
First, 
the length of average birth intervals have increased for all education groups, and
the increases are larger the higher the spell and the higher the education group.
Second, fertility has declined for all groups with larger declines the more educated the 
woman.
Across all groups, the likelihood of having an additional child depends strongly on the 
number of sons, with women with none or only one son having the highest probabilities.
Finally, as shown by the increasing sex ratios for women with no or only one son, there
has been a substantial increase in the use of sex selection.
As for fertility and birth intervals, the increases are larger the higher the spell and
with higher education.

The results show two very different spacing patterns in response to son preference.
At one extreme, rural women without education mostly follow the standard pattern of 
shorter spacing in the absence of sons.
The strength of this pattern does, however, depend critically on spell number.
For the second spell, although the average interval after a first-born girl is shorter
than after a first-born boy, the difference is small.
The differences in spacing are larger for the third spell and largest for the fourth.
For the fourth spell, the divergence in intervals has increased over time; the largest 
increases are for women with two or three boys, but there has also been an increase for 
women with only girls, likely because of an emerging use of sex selection.

At the other extreme, urban women with eight or more years of education have experienced 
an almost complete reversal of the traditional spacing patterns as the result of falling
fertility and the associated extensive use of sex selection.
Women with either no or one son now have substantially longer spacing than if they have 
two or more sons.
As shown by the increasingly significant deviations in the sex ratio from the natural sex 
ratio, these changes are due to increased use of sex selection rather than a sudden
reversal in son preference.
Only the second spell has not seen a reversal despite a substantially increased use of
sex selection, mainly because the average spacing with a first-born boy has increased 
independently by more than nine months.

The picture of how son preference affects spacing is, however, more complicated than 
excepted.
First, the differences in birth intervals for the first 50 percent of second births are 
negligible across sex compositions, despite these births clearly within the 
durations most likely to generate adverse outcomes.
Hence, most of the divergence in spacing patterns arise from longer spells.
Second, a side-effect of the extensive use of sex selection for third and fourth 
spells is that only-girls families are much less likely to have a very short birth 
interval, which may reduce the sibling competition for resources.

With a country as large as India, it is natural to expect variation in both fertility
behavior, spacing patterns, and the degree of son preference.
Although the regional results confirm this, the differences are less apparent than one would
expect given the prior research.
The ``North'' and the ``West'' regions have relatively similar spacing patterns, both
the result of rapidly falling fertility and substantial use of sex selection.
The exception is that the rural ``North'' still have higher fertility and
consequently less use of sex selection.
Even the ``South,'' otherwise considered to have the least son preference, has 
seen a substantial increase in the use of sex selection and associated changes 
in the spacing patterns.
Finally, fertility has declined fastest in the ``East'' region, but this has not 
translated into the use of sex selection.

The results here lead to three questions that future research should address.
First, what is behind the substantial increases in spacing even for women unlikely to use 
sex selection, and, related, why does spacing appear to increase with declining 
parity progression probabilities?
The distributional results suggest that most of the increases in average birth intervals 
come from the upper end rather than from a general increase across the distribution.
The longer average spacing and the relatively higher number of very long birth intervals 
seem to run counter to a general reduction in the length of women's reproductive spans 
found in earlier research \citep{Padmadas2004}.
Understanding whether the longer average spacing is by choice, because of unplanned 
pregnancies, or because of changes in economic circumstances that allow families to have 
an additional child later has important implications for both policy design and outcomes 
for mother and children.

Second, to what extent are the improvements in health for girls relative to boys the result 
of selection, the longer spacing between births, or changing son preference?
We know that health outcomes for girls appear to improve in the presence of sex selection 
\citep{Lin2014,Hu2015}.
Furthermore, sample selection can make it appear that girls are healthier, even though the 
underlying cause is a combination of sex selection and higher mortality together with 
recall error \citep{Portner2018a}. 
Understanding what role these factors play in better health for girls is especially 
important when evaluating policies that aim to directly limit the use of sex selection 
rather than changing preferences or incentives.
If the better health outcomes for girls are, for example, an unintended side-effect of the 
longer spacing that arises from sex-selective abortions, rather than because the smaller 
number of girls makes them more valued as is often assumed, then an effective ban on sex 
selection may, at least temporarily, worsen health outcomes for girls.

Finally, we need to understand how female labor force participation interacts with the use 
of sex selection.
Increased autonomy for women, arising, for example, from better opportunities for working 
outside the home, has been suggested as a way to increase women's status and thereby lower 
the use of sex selection \citep{Das-Gupta2016}.
This approach may, however, be a double-edged sword.
On the one hand, it is a clear benefit to the women who gain bargaining power, and it 
increases the cost of repeated sex-selective abortions because the increased duration 
between births would cause a stronger disruption in labor market participation.
On the other hand, it may further lower desired fertility, which, everything else equal, 
may lead to higher use of sex selection.
Understanding the trade-off between long-term benefits from improvements in women's labor 
force participation and short-term costs from potential increases in sex selection is of 
paramount importance.
